---
title: "Package miscset introduction"
author: "Sven E. Templer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{miscset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# About

A collection of miscellaneous methods to simplify various tasks,
including [plotting](#plot-methods), 
[data.frame and matrix](#data.frame-and-matrix-methods) transformations, 
[environment](#environment-methods) functions, 
[regular expression](#regular-expression-methods) methods, 
and [string and logical](#string-and-logical-methods) operations, 
as well as [numerical and statistical](#numeric-methods) tools. 
Most of the methods are simple but useful wrappers of common base `R` functions, 
which extend S3 generics or provide default values for important parameters.

# Installation

Install the package from `cran` or from `github` by using the following commands:

```{r eval=FALSE}
# from CRAN
install.packages("miscset")
# from github - latest version
devtools::install_github("setempler/miscset", build_vignettes = TRUE)
```

# Introduction

Development of the package is traceable at [github](https://github.com/setempler/miscset).
In case you find any bugs or have other issues concerning the package development,
feel free to make use of the github [issues](https://github.com/setempler/miscset/issues).

A more detailed help for each function can be viewed on the `R` help pages.
To view them, call the function's name with a `?` prepended, e.g.:

```r
?duplicates
```

The following chapters will describe the functions from the `miscset` package. The prerequisites to run all vignette examples are to load the following packages,
and generate the sample data:

```{r}
library(miscset)
library(ggplot2)
d <- data.frame(a=c(2,1,3,NA,1), b=2:6, c=5:1)
m <- matrix(letters[1:9], 3, 3, dimnames = list(1:3,1:3))
```

# Plot methods

([back to top](#about))

## ciplot

Plot a bargraph with error bars. Input data is a list with numeric vectors.
Functions to calculate bar heights (e.g. `mean` by default) and error bar sizes 
(e.g. `confint.numeric` by default) can be modified (e.g. `sd` for error bars).

```{r fig.height=3, fig.width=4, fig.align='center'}
ciplot(d)
```

## ggplotGrid, ggplotGridA4

Arrange ggplots on a grid (plot window or pdf file). Supply a list with `ggplot`
objects and define number of rows and/or columns. If a `path` is supplied, the
plot is written to that file instead of the internal graphics device.

```{r fig.height=2, fig.width=6, fig.align='center'}
ggplotGrid(list(
  ggplot(d, aes(x=b,y=-c,col=b)) + geom_line(),
  ggplot(d, aes(x=b,y=-c,shape=factor(b))) + geom_point()),
  ncol = 2)
```

## gghcl

Generate a character vector with html values from a color hue as in `ggplot`.

```{r comment=NA, fig.height=3, fig.width=4, fig.align='center'}
n <- length(d)
gghcl(n)
ciplot(d, col = gghcl(n))
```

## plotn

Create an empty plot. Useful to fill `layout`.

```{r fig.height=2, fig.width=2, fig.align='center'}
plotn()
```

# data.frame and matrix methods

([back to top](#about))

## sort

Sort `data.frame` objects. This extends the functionality of the base R distributed generic `sort`. Define multiple columns by column names as character vector or expression.

```{r, comment=NA}
d
sort(d, by = c("a", "c"))
```

## do.rbind

A wrapper function to row-bind `data.frame` objects in a list with `do.call` and `rbind`.
Object names from the list are inserted as additional column.

```{r comment=NA}
d[1:3,]
do.rbind(list(first=d[1:2,], second=d[1:3,]))
```

## enpaire

Generate a pairwise list (`data.frame`) of a matrix containing row and column id and upper and lower triangle values.

```{r comment=NA}
m
enpaire(m)
```

## squarematrix

Generate a symmetric (square) matrix from an unsymmetric one using column and row names.
Fills empty cells with `NA`.

```{r comment=NA}
m[-1,]
squarematrix(m[-1,])
```

## textable

Print a `data.frame` as latex table. Extends `xtable` by optionally including a latex header,
and if desired writing the output to a file directly and calling a system command to convert
it to a `.pdf` file, for example.

```{r comment=NA}
textable(d, caption = 'miscset vignette example data.frame', as.document = TRUE)
```

# Environment functions

([back to top](#about))

## help.index

Show the help index page of a package (with the list of all help pages of a package).

```r
help.index(miscset)
```

## lload

Load multiple R data objects into a list. List is of same length as number of files provided.
Sublists contain all respective objects. Simplification is possible if all names are unique.

```r
lload("folder/with/rdata/", "test*.RData")
```

## lsall

Return all current workspace (or any custom) object names, lengths, classes, modes and sizes in a `data.frame`.

```{r comment=NA}
lsall()
```

## rmall

Remove all objects from the current or custom environment.

```r
rmall()
```

# Regular expression methods

([back to top](#about))

## mgrepl

Search for multiple patterns in a character vector. Merge results by (custom) logical functions
(e.g. `any`, `all`) and use mutlicore support from the `parallel` package.
Optionally return the index (as with `which`).
Use `identity` to return a matrix with the results of each pattern per row.

```{r comment=NA}
mgrepl(c("a","b"), c("ab","ac","bc"), any)
mgrepl(c("a","b"), c("ab","ac","bc"), all)
mgrepl(c("a","b"), c("ab","ac","bc"), all, use.which = TRUE)
mgrepl(c("a","b"), c("ab","ac","bc"), identity)
```

## gregexprind

Retreive the `n`th or `"last"` index of an expression found in a character string.

```{r comment=NA}
gregexprind(c("a"), c("ababa","ab","xyz",NA), 1)
gregexprind(c("a"), c("ababa","ab","xyz",NA), 2)
gregexprind(c("a"), c("ababa","ab","xyz",NA), "last")
```

# String and logical methods

([back to top](#about))

## collapse

To collapse vectors, usually a call to `paste` or `paste0` setting the argument
`collapse` is applied. The collapse function is a wrapper of this functionality
applied to a single vector. It can be extended with the `.unique`, `.sort` and
`.decreasing` arguments, to return only unique and sorted values.

```{r comment=NA}
paste(letters, collapse = ";")
collapse(letters)
```

The `data.frame` method allows to collapse a data frame by identifier/grouping
columns (specified with `by`). Each group piece has then all value columns
collapsed with the default method.

In addition, the value columns can be collapsed to vectors, when `sep = NULL`
is selected, keeping a list of vectors for this column in the returned
data frame. `.sortby` allows to choose if the result should be sorted by
the grouping columns. `.unlist` provides a way to unlist value columns per
group, which is useful if the input has list columns.

```{r comment=NA}
# create example data
set.seed(12)
s <- s2 <- sample(LETTERS[1:4], 9, replace = TRUE)
s2[1:2] <- rev(s2[1:2])
d2 <- data.frame(group = rep(letters[c(3,1,2)], each = 3), 
                 value = s,
                 level = factor(s2),
                 stringsAsFactors = FALSE)

# collapse by all columns
collapse(d2)

# by first column, unique and decreasingly sorted values, groups sorted
collapse(d2, "/", 1, .unique = TRUE, .sort = TRUE, .decreasing = TRUE, .sortby = TRUE)

# by first column, but keep values as vectors
collapse(d2, NULL, 1)
```

The `data.frame` method also works on `data.table` objects, since it uses the methods
from the package of the same name to split the input into group pieces. If the input
inherits from `data.table`, the class is retained.

## leading0

Prepend `0` characters to numbers to generate equally sized strings.

```{r comment=NA}
leading0(c(9, 112, 5009))
```

## strextr

Split strings by a separator (`sep`) and extract all substrings matching a `pattern`.
Optionally allow multiple matches, and use multicore support from the `parallel` package.

```{r comment=NA}
strextr("xa,xb,xn,ya,yb", "n$", ",")
strextr("xa,xb,xn,ya,yb", "^x", ",", mult=T)
```

## strpart

Similar to `strextr`, but extracting substrings is done by setting an index value `n`.
Optionally roll the last value to `n` if it's index is less.

```{r comment=NA}
strpart("xa,xb,xn,ya,yb", ",", 3)
```

## strrev

Create reverse version of strings of a `character` vector.

```{r comment=NA}
strrev(c("olleH", "!dlroW"))
```

## duplicates, duplicatei

Determine duplicates. Return either a logical vector (`duplicates`) or an integer index
(`duplicatei`). Extends the base method `duplicated` by also returning `TRUE` for the
first occurence of a value.

```{r comment=NA}
data.frame(
  duplicate = d$a,
  ".d" = duplicated(d$a),
  ".s" = duplicates(d$a),
  ".i" = duplicatei(d$a))
```

# Numeric methods

([back to top](#about))

## p2star

Asign range symbols to values, e.g. convert p-values to significance characters.

```{r comment=NA}
p2star(c(0.003, 0.049, 0.092, 0.431))
```

## confint.numeric

Calculate confidence intervals. Extends the base method `confint` to numeric vectors.

```{r comment=NA}
d$a
confint(d$a, ret.attr = FALSE)
```

## ntri

Generate a series of triangular numbers of length `n` according to
[OEIS#A000217](https://oeis.org/A000217). The series for 12 rows of a triangle, 
for example, can be returned as in the following example.

```{r comment=NA}
ntri(12)
```

## scale0, scaler

Scale numeric vectors to a range of _0_ to _1_ with `scale0` or to a custom output range `r` and input range `b` with `scaler`.

```{r comment=NA}
d$c
scale0(d$c)
scaler(d$c, c(2, 6), b = c(1, 10))
```

## nunique

Return the amount (with `nunique`) or index (with `uniquei`) of unique values in a vector. Extends `plyr::nunique` by allowing `NA` values to be counted as a 'level'.

```{r comment=NA}
d$a
nunique(d$a)
nunique(d$a, FALSE)
uniquei(d$a)
uniquei(d$a, FALSE)
```

# Legal

([back to top](#about))

> Copyright (c) 2016 - Sven E. Templer
